LaMachine: A meta-distribution for NLP software

### 2019-03-29 docs/clarin_reviewer_response.md

We opted for this image to implicitly convey that LaMachine also has its uses in education. Showing actual development projects
would not be very meaningful to the uninformed reader.

## REVIEW 2

> The justification for such a system is well argued for but how such facility relates to other functions that CLARIN centres are
> supposed to serve is not entirely clear.

(see also the comments to reviewer 4 below, who I think has a similar question)

@@

## REVIEW 4

This reviewer poses some very good and interesting questions about the relation between LaMachine and the larger CLARIN
infrastructure.  It's hard to sufficiently address this in the paper in the limited space available, as we focus mostly
on providing an clear description on what LaMachine is (a distribution) and contains. We have now tried to emphasise
more the fact that LaMachine is a distribution, and how that relates to e.g. the CLARIN switchboard, as some of the
questions indicate we have not communicated this clearly enough.

I'd like to address each of the questions here:

@@

> The paper states "Support for macOS is limited because not all participating software supports it." Could it be a solution to containerize such
> software?

One of LaMachine's flavours provides a container solution (n.b: a single container for all selected software). But on
macOS containerisation often implies virtualisation, which in fact is also already provided by LaMachine as another
flavour. That indeed offers a viable way to use software in LaMachine that is not otherwise possible on macOS (or
Windows for that matter). The statement about limited support, however, applied to compiling software for macOS
natively. I added a small footnote to hopefully clarify this.

> Is LaMachine catalogued in the VLO, with a PID to the metadata?

No, nobody has ever requested this desire yet. I'll look into it. It may be a bit difficult as LaMachine is such a
meta-distribution that it is hard to pin down for any single scenario, but it would be interesting for the services
inside **a** LaMachine installation (our own public one perhaps).  I opened issue https://github.com/proycon/LaMachine/issues/138 to this end. It's not
in the paper as it does not offer sufficient space for such a discussion.

@@

> Does LaMachine aim at compatibility with the CLARIN Language Resource Switchboard? The paper states that LaMachine "transcends" the
> ambitions of the Switchboard, but more relevant is the question whether the two are compatible and can be integrated.

The "transcends" remark in our paper applies to the VRE project rather than to LaMachine, I now clarified it to prevent
confusion. LaMachine is not directly comparable with the switchboard as it offers no switchboard functionality. Now, the
switchboard itself could possibly be included in the LaMachine distribution. I briefly considered this but decided to
include a simpler portal instead as the switchboard did not meet all our demands and other initiatives were started to
compete with it (CLARIAH WP3 VRE). Tighter compatibility between the switchboard and services offered by **a** LaMachine
installation would be possible, but this is more a question for the switchboard developers than for us.

@@

In fact, the current switchboard already points to various webservices that are hosted in our LaMachine installation at
Radboud University, though this involved manual duplication of metadata rather than leveraging the codemeta we offer.

> Does LaMachine support and expose CMDI metadata for the data it processes? It seems that only CodeMeta is supported, which I do not
> believe to be a CLARIN standard. Why not CMDI?

LaMachine itself is a distribution so it doesn't process any data as such; that would be a question for particular tools
or frameworks within it, and would not be something we as a distribution would enforce.

### 2019-04-01 docs/lamachine_clarin2018.tex
typo and double ref in clarin paper

Within the scope of CLARIN, one might be inclined to draw comparisons with the CLARIN Language Resource Switchboard
\cite{zinn2016clarin}, which offers access to NLP tools through a web interface where users can upload their texts for
NLP processing, providing easy access and automated tool discovery. This, however, is notably different from what
LaMachine is. LaMachine is a software distribution; the CLARIN switchboard is not. Vice versa, LaMachine does not
currently provide any switchboard functionality, it merely offers a portal interface. The switchboard allows access to
tools hosted elsewhere, LaMachine installs and configures tools and only provides access to those it installed. Unlike
the switchboard, the audience for LaMachine is in the first place the data scientist and researcher who has
programming experience and wants to work with complex NLP software, on her own machine and unencumbered by high-level
interfaces if she so desires.

There are also other CLARIN initiatives that provide not just a switchboard, but aim at an eve more ambitious high-level virtual research
environment (VRE) with graphical interfaces to accommodate the researcher and allow execution and visualisation of
various NLP pipeline. The Dutch CLARIAH WP3 project\footnote{\url{https://github.com/meertensinstituut/clariah-wp3-vre}}
is such an initiative: LaMachine is currently developed in close collaboration with this project, in a role of technology supplier.

>>>>

Within the scope of CLARIN, one might be inclined to draw comparisons with the CLARIN Language Resource Switchboard
\cite{zinn2016}, which offers access to NLP tools through a web interface where users can upload their texts for
NLP processing, providing easy access and automated tool discovery. This, however, is notably different from what
LaMachine is. LaMachine is a software distribution; the CLARIN switchboard is not. Vice versa, LaMachine does not
currently provide any switchboard functionality, it merely offers a portal interface. The switchboard allows access to
tools hosted elsewhere, LaMachine installs and configures tools and only provides access to those it installed. Unlike
the switchboard, the audience for LaMachine is in the first place the data scientist and researcher who has
programming experience and wants to work with complex NLP software, on her own machine and unencumbered by high-level
interfaces if she so desires.

There are also other CLARIN initiatives that provide not just a switchboard, but aim at an even more ambitious high-level virtual research
environment (VRE) with graphical interfaces to accommodate the researcher and allow execution and visualisation of
various NLP pipeline. The Dutch CLARIAH WP3 project\footnote{\url{https://github.com/meertensinstituut/clariah-wp3-vre}}
is such an initiative: LaMachine is currently developed in close collaboration with this project, in a role of technology supplier.
